---
title: "hw2"
author: "shichenh"
date: "9/13/2017"
lab: "9-11am"
output: pdf_document
---

```{r}
suppressMessages(library(tidyverse))
knitr::opts_chunk$set(echo = F)
```

```{r load_data}
temp <- suppressMessages(read_csv("https://raw.githubusercontent.com/ucb-stat154/stat154-fall-2017/master/data/temperature.csv"))
colnames(temp)[1] <- "city"
temp.m <- temp[1:23,2:13]
# creating a df where city is the column names
city.temp <- data.frame(t(temp[,2:13]))
colnames(city.temp) <- unlist(temp[,1])

#variable name storing all cities
cities <- unlist(temp[,1])
```

# EDA1
```{r eda1}
#check the overall temperature of the cities
summary(city.temp)
```
Summary: The cities do not really have a super high max temperature, at the same tiem many
have mean temperature lower than 10(pretty cold XD)

```{r eda2}
table(temp$Area)
par(mfrow=c(3,3))
for (i in 1:length(cities)) {
  barplot(unlist(city.temp[,i]), 
        names.arg = 1:12,
        #unifying temperature limit to see the difference
        ylim = c(-10, 30),
        xlab = "month", ylab = "temperature(celsius)",
        main = (paste(cities[i], ":", paste(temp$Area[i], "region"))))
}
```
Seems like East region and north region tend to have temperatures during the winter. 
I wonder why not the West case? Maybe because the west region is closer to the ocean
so it makes the cities warmer?

# PCA
# raw 
```{r}
temp.m.mean <- scale(temp.m)
temp.cor <- cor(temp.m)

#eigen
temp.cor.eigen <- eigen(temp.cor)
#svd
temp.svd <- svd(temp.cor)
```

##1a displaying first 4 loadings
```{r}
loadings <- temp.svd$v
temp.svd$v[,1:4]
```


##1b obtaining principal components 
```{r}
pc <- temp.m.mean %*% temp.svd$v
pc[,]
```


## 1c display eigen value
```{r}
print("eigen values")
print(round(temp.svd$d, 2))
print("sum of eigen values")
sum(temp.svd$d)
```


# 2
## 2a
```{r}
sum_table <- data.frame(cbind(temp.svd$d, round(temp.svd$d/sum(temp.svd$d), 2), cumsum(temp.svd$d)/sum(temp.svd$d)))
colnames(sum_table) <- c("var", "percent", "percent.cumulative")
sum_table
```
## 2b
```{r}
barplot(sum_table$percent,
        names.arg = paste('PC', 1:length(temp.svd$d)),
        xlab = "Principal Component", ylab = "Percentage of Variance",
        main = "Scree Plot")
```

b.From the chart, it seems only two(probably3) PC contributes to significant amount
of variance. In other words, many data are redundant.  

c.I would choose the firs 3 dimensions, because it already contains 99% of the 
data variation.

#3
## 3a making a scatter plot
```{r}
temp.cities <- scale(temp[,2:13])
cities.pc <- data.frame(cbind(cities, temp.cities %*% loadings[,1:2], temp$Area))
#cities.pc <- data.frame(cbind(cities, temp.cities %*% t(pc[1:2,])))
colnames(cities.pc) <- c("city", "pc1", "pc2", "Area")
cities.pc$active <- c(rep(1, 23), rep(0, 12))
cities.pc <- cities.pc %>%
  mutate(active = factor(c(rep(1, 23), rep(0, 12)))) %>%
  mutate(pc1 = as.numeric(as.character(pc1)), pc2 = as.numeric(as.character(pc2)))

ggplot(cities.pc) +
  geom_point(aes(pc1, pc2, size = active, color = Area)) +
  geom_hline(yintercept=0) +
  geom_vline(xintercept=0) +
  geom_text(aes(pc1, pc2, label = city), hjust=0, vjust=2)
```

Most of the South Region cities are in the negative area of PC1, spread in pc2.
The North region cities are mostly in the positive axes of pc1. The West region
cities are mostly in the negative axes of city 2. Most of the East region cities 
are in the positive of both axes. 

## 3b square cosine 
```{r}
# Compute the square of the distance between an individual and the
# center of gravity
getdistance <- function(ind_row){
  return(sum(ind_row)^2)
  }
d2 <- apply(temp.m.mean, 1, (function(x) sum(x^2)))
# Compute the cos2
cos2 <- function(ind.coord, d2){return(ind.coord^2/d2)}
temp.cos2 <- round(apply(pc, 2, cos2, d2), 4)
temp.cos2[,1:4]
```

## 3 c
```{r}
ind.contrib <- data.frame(round(t(apply(pc,1, (function(x) 100/(nrow(pc) - 1) * x^2 /temp.svd$d))), 2))
rownames(ind.contrib) <- cities[1:23]
colnames(ind.contrib) <- paste('PC', 1:length(temp.svd$d))
ind.contrib
```
Contributing to  
PC1: Athens, Lisbon, Rome  
PC2: Reykjavik, Dublin, Moscow  

## 4a Correlation of all
```{r}
cor.all <- cor(temp[1:23,2:17], pc)
colnames(cor.all) <- paste0('PC', 1:12)
cor.all[,1:4]
```

## 4b correlation graph
```{r}
# Plot the correlation circle
a <- seq(0, 2*pi, length = 100)
plot( cos(a), sin(a), type = 'l', col="black",
      xlab = "PC1",  ylab = "PC2")
abline(h = 0, v = 0, lty = 2)
# Add active variables
arrows(0, 0, cor.all[1:12, 1], cor.all[1:12, 2], 
      length = 0.1, angle = 15, code = 2,
      col = "red")
arrows(0, 0, cor.all[13:16, 1], cor.all[13:16, 2], 
      length = 0.1, angle = 15, code = 2,
      col = "grey")
# Add labels
text(cor.all[1:12,], labels=rownames(cor.all)[1:12], cex = 1, adj=1, col = "red")
text(cor.all[13:16,], labels=rownames(cor.all)[13:16], cex = 1, adj=1, col = "grey")
```

## 4c
Most of the active components are negativedly correlated with PC1, half positive with
PC2 negative with PC2. While the sub varibles are positively correlated with PC1 
except 1.

# Final Conclusion
Despite not fully understanding how PCA works mathematically. From the graphs we can 
see it is very powerful in data compression or feature reduction.I am surprised how 
well the pc1, pc2 graph clusters cities from different regions.

