---
title: "lab8"
author: "shichenh"
date: "10/23/2017"
output: pdf_document
---

```{r}
library(ISLR)
library(FactoMineR)
library(ggplot2)
library(dplyr)
```

```{r}
Default <- Default
```

# EDA
```{r}
ggplot(Default) +
  geom_point(aes(x=balance, y=income, color=default)) +
  labs(title="Scatter Plot of Balance and Income")


ggplot(Default) +
  geom_density(aes(x=balance, color=default)) +
  labs(title="Kernel Density of Balance")

ggplot(Default) +
  geom_density(aes(x=income, color=default)) +
  labs(title="Kernel Density of Income")
```


# Default
## OLS Regression
```{r}
default_numeric <- rep(0, nrow(Default))
default_numeric[Default$default == 'Yes'] <- 1
Default$default_num <- default_numeric
ols_reg <- lm(default_num ~ balance, data = Default)
summary(ols_reg)
```

## Logistic Regression 
```{r}
logreg_balance <- glm(default ~ balance, family = binomial, data = Default)
summary(logreg_balance)$coefficients
```

```{r regressing_balance}
balance <- seq(100, 2000, by=100)
odds.balance <- exp(predict(logreg_balance, newdata = data.frame(balance)))
pred.balance <- odds.balance/(1+odds.balance)
plot(pred.balance)
```
```{r regressing_student}
logreg_student <- glm(default ~ student, family = binomial, data = Default)
summary(logreg_student)$coefficients
```

If the person is a student, then the log odds of it being default increases by 0.4048871.

```{r}
logreg_all <- glm(default ~., family = binomial, data = Default)
summary(logreg_all)
```

Only *income* is not statistically significant.  
The apparent contradiction between the opposite signs of the
student coefficients may due to colinearity among student, balance and income.  

# Stock Market
```{r}
Smarket <- Smarket
```

## Variables
```{r}
cor(Smarket %>% select(-Direction))
#pca
PCA(Smarket[ ,-9])
#
```

The lag variables are not so correlated with today's return. So previous days return does not seem to correlate with today's return.

```{r}
ggplot(Smarket) +
  geom_point(aes(x=Year, y=Volume))
```

## Logistic Regression
```{r}
logreg_smarket <- glm(Direction~., family = binomial, data = Smarket %>% select(-c(Year, Today)))
summary(logreg_smarket)
```

None of the coefficients are significant. Lag1 coefficient is -0.073074. The negative sign means the Lag1 is inversely proportion to the log odds of Ups/Downs. If log1 increases, today's return is less likely to be up.

```{r}
pred.smarket <- predict(logreg_smarket, newdata= Smarket %>% select(-c(Year, Today, Direction)), type="response")
pred.smarket[1:10]
```


# Parameters Estimation
## Newton-Raphson
```{r}
x <- model.matrix(Direction~., data = Smarket %>% select(-c(Year, Today)))

y <- as.matrix(Smarket %>% 
  mutate(Direction = ifelse(as.numeric(Direction) == 2,1,0)) %>% select(Direction)
)
```

```{r}
#number of iterations 
n.iter = 100
b0 <- as.matrix(rep(0, 7))
i <- 1

p <- exp(x %*% b0)/(1 + exp(x %*% b0))
w <- diag(p[1:length(p)]) * (1-p[1:length(p)])
z <- x %*% b0 + solve(w) %*% (y - p)
b <- solve(t(x) %*% w %*% x) %*% t(x) %*% (w) %*% z

while (sum(abs(b - b0)) >= 0.1^5 && i <= n.iter) {
  b0 <- b
  p <- exp(x %*% b0)/(1 + exp(x %*% b0))
  w <- diag(p[1:length(p)]) * (1-p[1:length(p)])
  z <- x %*% b0 + solve(w) %*% (y - p)
  b <- solve(t(x) %*% w %*% x) %*% t(x) %*% (w) %*% z
  i <- i + 1
}

print(sum(abs(b-b0)))
print(b)
```

## Simplified Algoirthm
```{r}
n.iter = 100
b0 <- as.matrix(rep(0, 7))
i <- 1
```

```{r}
p <- exp(x %*% b0)/(1 + exp(x %*% b0))
x.hat <- t(x) %*% (p*(1-p))
b <- b0 + solve(t(x) %*% x) %*% t(x) %*% (y-p)

while (sum(abs(b-b0)) >= 0.1^8 & i <= n.iter) {
  b0 <- b
  p <- exp(x %*% b0)/(1 + exp(x %*% b0))
  x.hat <- x
  for (j in 1:length(p)) {
    x.hat[i,] = x.hat[i,] * p[i] * (1 - p[i])
  }
  b <- b0 + solve(t(x) %*% x.hat) %*% t(x) %*% (y-p)
  i <- i + 1
}

print(sum(abs(b-b0)))
print(b)
```

